\documentclass[oneside,final,12pt,a4paper]{extreport}
\renewcommand{\thetable}{\Roman{table}}\usepackage{hyperref}
\usepackage{hyperref}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    citecolor=cyan,
    pdftitle={Overleaf Example},
    pdfpagemode=FullScreen,
    }
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[affil-it]{authblk}
\usepackage[english]{babel}
\usepackage[backend=biber,style=ieee,autocite=inline]{biblatex}
\usepackage{blindtext}

\usepackage{float}
\usepackage{amsmath,amsfonts}
\usepackage{url}
\usepackage{titlesec}
\usepackage{csquotes}
\usepackage{longtable}
\titleformat{\section}[hang]{\fontsize{20}{24}\selectfont\filcenter}{\Roman{section}}{1em}{}
\titleformat{\subsection}[hang]{\itshape}{\Alph{subsection}.}{1em}{}[]
\titleformat{\subsubsection}[runin]{\itshape}{\arabic{subsubsection})}{1em}{}[$:$]
\titlespacing{\subsubsection}{1em}{1em}{1em}
\titleformat{\paragraph}[runin]{\itshape}{\alph{paragraph})}{1em}{}[$:$\quad]
\titlespacing{\paragraph}{2em}{1em}{1em}
\usepackage{vmargin}
\setpapersize{A4}
\setmarginsrb{2.5cm}{2cm}{2cm}{2cm}{0pt}{10mm}{0pt}{13mm}
\usepackage{setspace}
\sloppy
\setstretch{1.5}
\usepackage{indentfirst}
\usepackage{pgfplots}
\parindent=1.25cm
\usepackage{caption}
\usepackage{subcaption}
\DeclareCaptionLabelSeparator{figSep}{.\quad}
\title{CRISP-DM Report: BiznesDengi}
\author{Aleliya Turushkina, Arina Petuhova, Maksim Ilin, Ivan Ilyichev, Lev Permiakov}


\begin{document}

\maketitle

\newpage

\section{Data Understanding}
\subsection{Data Selection}
Before collecting any data, we decided which stocks, time periods, and tables to use based on the business goal:

The dataset includes ordinary shares of the 249 most traded Russian companies on the Moscow Exchange. The number 249 comes from the Broad Market Index (IMOEX) plus the most liquid securities from the second and third trading levels, as defined by the Exchange. This number balances variety and data quality. Fewer stocks, such as the top 50 only, would limit choices and increase risk from relying on too few companies. More stocks, such as all 800 available, would include many rarely traded instruments with too many missing values and unreliable prices. We selected only ordinary shares because they give voting rights and dividends, which match our investment goal. We excluded preferred shares and depositary receipts to avoid differences in how dividends work and how often they are traded.

The collected data covers 16 years, from 1 February 2010 to 1 February 2026. We chose this period because it includes different market situations: the 2014 sanctions, the 2020 pandemic crash, and the 2022 volatility. The last year (2025–2026) is kept aside for testing on unseen data. A shorter period would risk the model learning only one type of market condition. A longer period would include price behavior from before 2008 that no longer reflects today's market. Therefore, we excluded all data before 2010.

We selected four types of tables. The Market Candles tables give price data at six intervals: 10 minutes, 1 hour, 1 day, 1 week, 1 month, and 1 quarter. Short intervals help us understand price changes within a day and choose good entry points. Longer intervals show clear trends over weeks and months. The Market History table provides daily open, high, low, close, and trading volume. This data is needed to calculate returns and losses over the one-year period. For both Market Candles and Market History, we included data from all trading modes because limiting to a single mode would exclude many trades and reduce data completeness. The Dividends table records past dividend payments per share, including the date when shareholders must be registered and the currency used. This allows us to calculate total return, not just price change. The Security Descriptions table gives fixed information about each stock, such as industry sector, listing level, and high-risk flags. This helps us filter stocks by company characteristics.

\subsection{Data Collection}

The dataset was collected by parsing historical market data from the Moscow Exchange Information and Statistical Service (MOEX ISS) API. The data collection was implemented using the \texttt{apimoex} Python library, which is convenient for accessing MOEX market data endpoints. The collected data covers the period from \texttt{01.02.2010} to \texttt{01.02.2026} and includes daily trading records, historical dividend payments, candlestick data across multiple time intervals, and securities description for the predefined list of 249 most popular Russian stocks.

\subsection{Dataset Description}
The dataset is organized into four table types:

\subsubsection{Market Candles Tables}
These tables represent historical candlestick data aggregated across multiple time intervals for all securities and trading boards. The data is stored in a hierarchical folder structure, with separate subfolders for each interval: \texttt{10min}, \texttt{1hour}, \texttt{1day}, \texttt{1week}, \texttt{1month}, and \texttt{1quarter}. Within each interval folder, individual CSV files follow the naming pattern \texttt{\{ticker\}-\{interval\}.csv}. The CSV schema includes the candle start time (\texttt{begin}), the four standard price points (\texttt{open}, \texttt{high}, \texttt{low}, \texttt{close}), the total trading value in rubles (\texttt{value}), and the volume in lots (\texttt{volume}).

\subsubsection{Dividends Table (\texttt{dividends.csv})}
This table captures the historical dividend payments for securities in the dataset. Each record contains the ticker (\texttt{secid}), the international identifier (\texttt{isin}), the critical register closure date (\texttt{registryclosedate}), the dividend amount per share (\texttt{value}), and the currency of payment (\texttt{currencyid}).

\subsubsection{Market History Table (\texttt{market\_history.csv})}
This table contains daily trading records for all securities across all trading boards. Each row is uniquely identified by the composite key of ticker (\texttt{SECID}), trading board identifier (\texttt{BOARDID}), and trading date (\texttt{TRADEDATE}). The CSV schema includes opening price (\texttt{OPEN}), intraday peak (\texttt{HIGH}), intraday trough (\texttt{LOW}), and closing price (\texttt{CLOSE}), along with trading volume in lots (\texttt{VOLUME}) and total monetary value of all trades (\texttt{VALUE}). Additional metrics include the volume-weighted average price (\texttt{WAPRICE}), the number of individual transactions (\texttt{NUMTRADES}), the denomination currency (\texttt{CURRENCYID}), security's short name (\texttt{SHORTNAME}), official settlement closing price (\texttt{LEGALCLOSEPRICE}), trend closing price (\texttt{TRENDCLSPR}), market prices (\texttt{MARKETPRICE2}, \texttt{MARKETPRICE3}), admitted quote (\texttt{ADMITTEDQUOTE}), associated trade values (\texttt{MP2VALTRD}, \texttt{MARKETPRICE3TRADESVALUE}, \texttt{ADMITTEDVALUE}), weighted average value (\texttt{WAVAL}), trading session identifier (\texttt{TRADINGSESSION}), and trade session date (\texttt{TRADE\_SESSION\_DATE}).

\subsubsection{Security Descriptions Table (\texttt{security\_descriptions.csv})}
This table acts as a static reference, containing one row per ticker with unchanging metadata about each security. Key columns include the ticker (\texttt{SECID}), the issuer's full name (\texttt{NAME}), the trading short name (\texttt{SHORTNAME}), the international identifier (\texttt{ISIN}), the state registration number (\texttt{REGNUMBER}), the nominal value (\texttt{FACEVALUE}) in the specified unit (\texttt{FACEUNIT}), the total issue size (\texttt{ISSUESIZE}), the issue date (\texttt{ISSUEDATE}), the listing level (\texttt{LISTLEVEL}), and classification fields (\texttt{GROUP}, \texttt{TYPE}, \texttt{TYPENAME}). Additional metadata covers the emitter identifier (\texttt{EMITTER\_ID}), the registry date (\texttt{REGISTRY\_DATE}), trading session flags (\texttt{MORNINGSESSION}, \texttt{EVENINGSESSION}, \texttt{WEEKENDSESSION}), qualified investor status (\texttt{ISQUALIFIEDINVESTORS}, \texttt{QUALINVESTORGROUP}), high-risk flag (\texttt{HIGHRISK}), and depositary organization details (\texttt{DEPOSITARY\_ORG\_NAME}, \texttt{DEPOSITARYNOTETYPE}).

\end{document}
